### 2019-08-22 paper-jgc.tex
Title, abstract, intro & conclusions

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
%\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
%gsave
%newpath
%  20 20 moveto
%  20 220 lineto
%  220 220 lineto
%  220 20 lineto
%closepath
%2 setlinewidth
%gsave
%  .4 setgray fill
%grestore
%stroke
%grestore
%\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{array}
\usepackage{subfigure}
\usepackage{adjustbox}
\usepackage{listings}
\usepackage{url}
\usepackage{cite}

%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

%\title{Evolution of Processes in Delivering Reliable Software within the Distributed European Research Infrastructures%\thanks{Grants or other notes
%%about the article that should go on the front page should be
%%placed here. General acknowledgments should be placed at the end of the article.}
%}
\title{Delivering more Reliable Software within the INDIGO-DataCloud project: a converging evolution of software engineering practices to support European Research e-Infrastructures}
%\subtitle{}

%\titlerunning{Short form of title}        % if too long for running head

\author{P. Orviz Fernandez \and
        M. David \and
        D. C. Duma \and
        E. Ronchieri \and
        J. Gomes \and
        D. Salomoni
%
%First Author         \and
%        Second Author %etc.
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{
P. Orviz Fernandez \at
CSIC, Santander, Spain \\
\email{orviz@ifca.unican.es}
\and
M. David \at
LIP, Lisbon, Portugal \\
%Tel.: +123-45-678910\\
%Fax: +123-45-678910\\
\email{david@lip.pt}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
\and
D. C. Duma \at
INFN - CNAF, Bologna, Italy\\
\email{cristina.aiftimiei@cnaf.infn.it}
\and
J. Gomes \at
LIP, Lisbon, Portugal \\
%Tel.: +123-45-678910\\
%Fax: +123-45-678910\\
\email{jorge@lip.pt}
\and
E. Ronchieri \at
INFN - CNAF, Bologna, Italy\\
\email{elisabetta.ronchieri@cnaf.infn.it}
\and
D. Salomoni \at
INFN - CNAF, Bologna, Italy\\
\email{davide.salomoni@cnaf.infn.it}
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
From the advent of Grid technology -- as the new paradigm of distributed
computing -- to the current days of Cloud computing models, the continuous need
of new tools and services to match the scientific community requirements has been
addressed in Europe through dedicated software development projects for
e--Infrastructure creation, operation and management.
This work presents the most significant software quality breakthroughs obtained in
one of such projects, {\sl INDIGO--DataCloud}, the main challenges and barriers 
confronted throughout the lifespan of the project, and how they were partially or 
totally overcome.
The knowledge base established throughout the last 15 years 
of diverse software development initiatives in Europe for sustaining distributed research 
e-Infrastructures, supported by the advances in the area of software engineering, definitely
%constituted a significant breakthrough in the reliability of the software delivered in the 
contributed to improve the reliability of the software delivered, and consequently, the 
operational stability of the European e--Infrastructures. {\sl INDIGO--DataCloud} project is 
a good evidence of such insights, where, unlike the preceding trend found in past projects,
the enforcement of Software Quality Assurance practices has been present since the very early 
stages of the software lifecycle.
\keywords{Software Reliability \and Quality Assurance \and Software Metrics \and Software Testing
Techniques \and DevOps}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{intro}

Throughout the last 15 years, Europe has have continuously invested on software development initiatives to provide
research e--Infrastructures with the capability of supporting unified access to large--scale computing and intense 
data analysis facilities. The developed tools have enabled the federation of geographically distributed resource 
providers, with the ultimate aim of supporting the growing needs of the scientific comumunities. According to their
requirements, those research e--Infrastructures have been evolving, exploiting diverse distributed computing 
technologies namely Grid, in the first place, and, eventually, Cloud.

The urging needs of providing stable e--Infrastructures, which were based on novel distributed computing technologies, 
initially drove the efforts spent in software development towards the provision of new functionalities, relegating the 
adoption of software engineering methodologies to a second place. The released software was less tested, resulting in 
a high rate of rollbacks and patches applied to production systems, when compared to the subsequent projects. 
Thereby, the stability and reliability of the e--Infrastructures were often compromised \cite{lingrand-egee}, leading 
to outages in the participant resource centers. Back in the days when Grid middleware was actively developed, 
researchers reported the need of improvements both in terms of reliability and performance \cite{campana-atlas}, 
eventually developing their own custom solutions \cite{kindermann}\cite{mendez-lorenzo} on top of the existing services.

In an effort to improve the operational experience of research communities, it soon became apparent the need for a better 
balance between the addition of new features and the reliability of the released software. Consequently, incoming European 
funding programmes and open calls progressivelly stressed the fact of devoting a more substantial part of the software 
lifecycle to the improvement of the quality and sustainability of the products being delivered. At this point,
evolving software engineering practices were considered and gradually adopted in order to define and implement the quality
procedures, the organization of cross--functional teams and the deployment of pilot testbeds.

The increasing prominence of software quality procedures observed in the analysis of the last decade of EU--funded
projects related to the development of distributed computing solutions --as described in Section~\ref{sec:ev}--, is 
pictured in the software project herein introduced: {\sl INDIGO--DataCloud}. As we will thoroughly describe in the next
sections, the ground base laid by these preceding projects and the adoption of practices taken from the prevailing 
software engineering methodologies --in particular the Agile Software Development \cite{agile-manifesto} and the DevOps
culture \cite{zhu}--, were the catalyst to establish a set of Software Quality Assurance (SQA) criteria 
-- as explained in Section~\ref{software-quality-procedures} -- that guided the software lifecycle of the software 
produced.

Following the DevOps principles, the SQA criteria requires the verification of the source code on each change in the 
production version, defining the essential requirements to be processed. Such scenario is hardly accomplished without
the aid of automation, and as a consequence -- as described in Section~\ref{software-automation} -- the testing and 
validation solution adopted by {\sl INDIGO--DataCloud} relied on the great advancements that continuous integration 
(CI) tools have undergone in the last years. Accordingly, both the validation and delivery of {\sl INDIGO--DataCloud} 
software was tackled by automation means.

The fundamental shift brought by the INDIGO--DC SQA procedures was 
to act at the early stages of the software development lifecycle. Thus, instead of validating the software product once 
it was packaged within a pre-release stage, the source code itself is verified for each minor change. This strategy 
allows to improve the sustainability of the software, by checking the compliance with code style standards, and its 
reliability, through the execution of functional and integration tests, ensuring the total absence of regressions. The results
presented in this paper show a divergence between the high rate of bug fixes resolved at development time and the 
significantly lower number of bugs detected by the end users. Consequently, bugs are more effortessly fixed since they 
are early captured and end users interface with more reliable and stable software.

The paper is organized as follows. Section~\ref{sec:b} summarizes the essential context of the research problem.
Section~\ref{sec:ev} details the evolution of research on software reliability in the selected EC--projects. 
Section~\ref{sec:ntsr} provides information about new research trends in software reliability based upon the
experiences from the recent {\sl INDIGO--DataCloud} project (INDIGO--DC). Conclusions are drawn in Section~\ref{sec:con}.

%%%%%%%%%%%%%%%%%%%%%%%%

%In this paper, we outline the set of EC--funded projects, ranging from 2001 to nowadays,
%that have raised the awareness and, at the same time, laid the ground work for the establishment of 
%thorough SQA practices to guide the software development lifecycle within the {\sl INDIGO--DataCloud}
%project. 
%%
%%For each project, the challenges, the methodologies and the achievements towards software 
%%reliability are described.
%%During the last 15 years the provision of European e--Infrastructures,
%%supporting unified access to large--scale computing and intense data analysis
%%applications have been driven by the needs of scientific communities as part of their
%%pan--European research collaborations. 
%%One such example is the Worldwide
%%Large Hadron Collider (LHC) Computing Grid infrastructure, driven
%%by the particle physics scientific community, built to satisfy the computing needs
%%of processing, storing and analysis of the very large amount of data produced by the
%%LHC particle accelerator.
%
%A significant number of the innovation projects and technology transfer
%activities presented in Section~\ref{sec:ev}
%were funded by the 5$^{th}$, 6$^{th}$ and 7$^{th}$ Framework Programmes of the European Commission (EC)
%and recently by the Horizon 2020 programme \cite{h2020}. Their main goal was to develop
%software solutions that provide a seamless and transparent access by researchers to
%distributed computing and data resources.
%
%Major developments carried throughout the years were focused on providing new
%functionalities, whereas in the early days the adoption of software engineering methodologies had a lower priority.
%The released software was less tested, resulting in a higher rate of
%rollbacks and patches applied to production systems, compared to the later projects. Thereby, the stability and
%reliability of the e--Infrastructures was often compromised \cite{lingrand-egee}, leading to outages
%in the participant resource centers. Back in the days when Grid middleware was actively developed, researchers
%reported the need of improvements both in terms of reliability and performance \cite{campana-atlas}, eventually
%developing their own custom solutions \cite{kindermann}\cite{mendez-lorenzo} on top of the existing services.
%
%In an effort to improve the research community's experience, it became apparent the need for a better
%balance between the addition of new features and the reliability of the released software. Consequently,
%incoming European funding programmes and open calls stressed the fact of devoting a substantial part of the software lifecycle to the improvement of the quality of the software being delivered. At this point, evolving software engineering practices were considered
%and gradually adopted to define and implement quality procedures, the organization of
%cross--functional teams and deployment of pilot testbeds.
%
%As a result of the analysis the aforementioned projects, 
%one can observe a continuous increase in the prominence and robustness of software
%testing and validation procedures \cite{aiftimiei}. The
%evolution of the software engineering methodologies (creation of the Agile
%Manifesto \cite{agile-manifesto}, and rise of DevOps practices \cite{zhu}) together
%with the parallel advancements in the information and communication technology
%field (such as virtualization, the enabling technology for Cloud computing and
%operating--system--level virtualization known as containerization \cite{soltesz}) motivated this trend.
%This evolution was complemented by the emergence of automation and event--response tools.
%
%The former know-how combined with the continuous advancements in sofware engineering 
%have been the catalysts for drawing attention to the relevance of SQA ..

\section{Background}
\label{sec:b}

In this section, we detail the level of knowledge requested in delivering reliable software in order
to understand the effort employed in the EC-funded software projects over time. We divide the literature
into four different domains that intersect this work: 1) software requirements analysis, 2) software
coding,
3) metrics and measurements, 4) testing.

\emph{Software requirements analysis.} They are defined at various stages of a software development process
as a specification of what should be implemented. According to IEEE \cite{radatz}, requirement
analysis is a process for studying user needs to arrive at a definition of system, hardware or software
requirements. Characteristics of requirements that can assess software reliability are consistency and
correctness \cite{sommerville}: the former aims at providing predictable, maintainable and reliable
results, and the latter usually refers to the satisfaction of certain goals.

\emph{Software coding.} According to IEEE \cite{radatz}, coding is the process of expressing a computer
program in a programming language. To meet specific requirements, a program can be developed by using
high-level or low-level programming languages that envelope different characteristics and may affect the
result and quality of software. Furthermore, software is generally written by geographically distributed
teams whose developers can change over time, therefore it is extremely important for source code to be
understandable \cite{tashtoush}. There are various instruments that can be adopted during the software
development process to easily adapt software: programming features that influence the
code readability and maintainability \cite{buse} and standardized code that promotes the comparability of results \cite{baggen}.

\emph{Metrics and measurements.} According to IEEE \cite{radatz}, metrics is a function, with input as
the software data, and output is a value which could decide on how the given attribute affects the software.
Metrics and measurements have an important role in the development process \cite{fenton, capers}. Metrics
are used to improve the ability to control, identify and measure the essential parameters of software
during its development, leading to improve the results of the software product \cite{goodman}. With a correct
usage of metrics, it is possible to determine which parts of the software require to be changed or modified.
In the literature, different types of metrics exist to be able to provide information on various software
aspects, such as complexity \cite{mccabe} and object orientation \cite{chidamber, lorenz}. Often they have
similar definitions, so as a consequence they can be implemented in different ways according to the
interpretation of their definition. This is the reason why it has been always a challenge to select the
proper software metric tools.

\emph{Testing.} According to IEEE \cite{radatz}, testing is a process of operating a system or component
under specified conditions, observing or recording the results, and making an evaluation of some aspect of
the system or component. Testing is an important factor for software quality and reliability improvement. This
activity is quite complicated due to the variety of programming languages, operating systems and hardware
platforms that evolve over time. In the current literature, there are many types of software
testing methods \cite{myers} that aim at finding errors, bugs and faults during the software development
process, possibly in the early stages. Over time, testing metrics have been defined to assess software
quality \cite{athanasiou}: code coverage is the most frequently used metric for testing software reliability
and there exist many tools for dynamic code coverage estimation according to the programming language
\cite{horgan}. A dedicated effort has to be taken into account in order to integrate testing solutions
during the software development process.

\section{Evolution of research on software reliability in EC--projects}
\label{sec:ev}

The EC projects considered in this paper are listed chronologically in Table~\ref{tab:eup}.
Each project paved the way for the following ones, according to a logic evolution in which
the role of the quality and reliability of the software has
consistently increased. As shown in Table~\ref{tab:feat}, the most recent projects are more
committed to, and aware of, software reliability practices.
Table~\ref{tab:feat} also shows the topics addressed in the following sections, highlighting
the commonalities and specific features on software reliability.

\begin{table}[!h]
\renewcommand{\arraystretch}{1.3}
\caption{List of EC--projects}
\label{tab:eup}
\centering
\begin{tabular}{p{1.6cm}p{2cm}p{5cm}l}
\hline
\hline
\\
Logo & Short Name & Long Name & Duration\\
\hline
\hline
\begin{minipage}{.3\textwidth}
\includegraphics[width=15mm,height=7.5mm]{images/datagrid}
\end{minipage}
    & DataGrid &
Research and Technological Development for an International Data Grid & 2001--2003\\
\begin{minipage}{.3\textwidth}
\includegraphics[width=15mm,height=7.5mm]{images/egee}
\end{minipage}
     & EGEE I, II, III &
Enabling Grids for E--sciencE & 2004--2010\\
\begin{minipage}{.3\textwidth}
\includegraphics[width=15mm,height=7.5mm]{images/etics}
\end{minipage}
     & ETICS 1, 2 &
E--Infrastructure for Testing, Integration and Configuration of Software & 2006-2010\\
\begin{minipage}{.3\textwidth}
\includegraphics[width=15mm,height=7.5mm]{images/emi}
\end{minipage}
     & EMI &
Europeean Middleware Initiative & 2010--2013\\
\begin{minipage}{.3\textwidth}
\includegraphics[width=15mm,height=7.5mm]{images/egi-inspire}
\end{minipage}
     & EGI--Inspire &
Integrated Sustainable Pan--European Infrastructure for Researchers in Europe
 & 2010--2014\\
\begin{minipage}{.3\textwidth}
\includegraphics[width=15mm,height=7.5mm]{images/egi_engage}
\end{minipage}
     & EGI Engage &
Engaging the EGI Community towards an Open Science Commons
 & 2015--2017\\
\begin{minipage}{.3\textwidth}
\includegraphics[width=15mm,height=7.5mm]{images/indigo}
\end{minipage}
     & INDIGO--DataCloud &
INtegrating Distributed data Infrastructures for Global ExplOitation
 & 2015--2017\\
\hline
\hline
\end{tabular}
\end{table}

\begin{table*}[!h]
\renewcommand{\arraystretch}{1.3}
\caption{Features on software reliability in EC--projects}
\label{tab:feat}
\centering
\begin{tabular}{p{2.6cm}lllllll}
\hline
\hline
Feature & DataGrid & EGEEs & ETICSs & EMI & EGIs & INDIGO--DC\\
\hline
\hline
Architecture Task Force            &$\surd$&$\surd$&$\surd$&$\surd$&$\surd$&$\surd$\\
Communication Handling             &$\surd$&$\surd$&$\surd$&$\surd$&$\surd$&$\surd$\\
Requirements Handling              &$\surd$&$\surd$&$\surd$&$\surd$&$\surd$&$\surd$\\
Agile Methodologies                &       &$\surd$&$\surd$&$\surd$&$\surd$&$\surd$\\
Source code inspection             &       &       &       &$\surd$&       &$\surd$\\
Build/Testing Management Procedure &$\surd$&$\surd$&$\surd$&$\surd$&       &$\surd$\\
Software Product Metrics           &$\surd$&$\surd$&$\surd$&$\surd$&$\surd$&$\surd$\\
Quality Criteria Definition        &$\surd$&       &$\surd$&$\surd$&$\surd$&$\surd$\\
Automatic Certification            &       &       &$\surd$&       &$\surd$&$\surd$\\
Auto--generated Documentation      &       &$\surd$&$\surd$&$\surd$&$\surd$&$\surd$\\
DevOps practices adoption (CI, CD) &       &       &       &       &       &$\surd$\\
\hline
\hline
\end{tabular}
\end{table*}

\subsection{DataGrid}

The {\sl DataGrid} \cite{cordis:datagrid} project (Jan 2001 -- Dec 2003)
main goal was to provide scientific communities (including physics, biology and
earth sciences) with intensive computing and large--scale dataset analysis capabilities.
The project brought together 21 academic and industry partners, from 15 different countries \cite{gagliardi}.
Grid was the emerging technology to be used in order to address their requirements,
thus the project delivered its own software distribution, named EDG (EU {\sl DataGrid}), strongly
based on Globus middleware components and services \cite{globus}. The project was organized in 12 work packages, from
which 5 were devoted to software development and coordinated by an
Architecture Task Force, supervising the overall design and technical consistency
of the developments.

Without previous experience in such widely
collaborative projects, a big effort was spent in devising solutions for several
challenges \cite{datagrid}, namely 1) the communication overhead (as a
result of the large geographical separation of the involved parties in the
development tasks), 2) the fast evolution of the requirements from the user
communities (50 use cases from the three scientific area), and 3) the lack of a body
of knowledge for academic software engineering.

The Agile manifesto \cite{agile-manifesto} was by then an emerging methodology, not yet
considered by the project. As such, the project's
development design suffered from the lack of the methods, tools, techniques and best
practices emerging from this new discipline of software engineering
\cite{agile}. Nevertheless, the project focused on the experiences and procedures
used by diverse collaborative open source software projects, such as Linux and Apache, trying to
achieve a higher maturity level \cite{cmm}.

Project monitoring and reporting mechanisms allowed to assess the risks, efforts and documentation
produced. Development guidelines \cite{datagridguide} were produced, covering the various phases, including
1) packaging, 2) test and validation, and 3) style and naming convention. The specification of quality indicators
\cite{datagridindicators} facilitated the detection of deviations in the collected metric data. Figure
\ref{fig:efficiency} shows the efficiency performance indicator, measured in production environments as the
number of jobs successfully completed divided by the total number of jobs submitted. As it can be observed,
there is a positive trend of efficiency over the months, with the unique exception of the LCG1 experiment.
\begin{figure}[h]
\centering
%\includegraphics[width=0.8\textwidth, height=50mm]{images/datagridEff.png}
\includegraphics[width=1\textwidth]{images/datagridEff.png}
\caption{Efficiency \% = Number of jobs successfully completed / Total Number of jobs submitted: results are provided for five experiments involved in the project.}
\label{fig:efficiency}
\end{figure}

\subsection{Enabling Grids for E--sciencE}

The three phases of Enabling Grids for E--sciencE ({\sl EGEE}, Apr 2004 -- Apr 2010)
\cite{cordis:egee,cordis:egee2,cordis:egee3} projects brought together
scientists and engineers from over 240 institutions in 45 countries aiming to
provide a seamless Grid infrastructure for e--Science. {\sl EGEE--II} and {\sl EGEE--III}
featured the internationalization (outside Europe) of the project, embracing worldwide research
institutions and user communities~\cite{egee}. The software to sustain the increasing
requirements coming from the diverse scientific communities needed developing a
rich set of new services, while maintaining a sustainable infrastructure for
Grid computing. This e-Infrastructure was eventually used by more than 15 thousand researchers
and deployed in over 250 institutions.

The {\sl gLite} middleware \cite{glite} was the
official software distribution of {\sl EGEE} as of 2006, after two years of prototyping and
re--engineering efforts to converge with the {\sl LHC} Computing Grid (LCG--2), Virtual
Data Toolkit (VDT) and Condor \cite{condor} software stacks. The
development team was comprised of more than 80 people from 12 academic and
industrial partners, which issued more than 10 thousand bug fixes, 1.7 thousand patches and
defined over 300 development tasks tracked using bug/task management tools.

The source code was available at a private, centralized version control system.
The code passed through a manual certification procedure at the time of the release.
This procedure aimed to improve the reliability of the software components by applying
acceptance criteria checks at the pre--release stage
\cite{egee:acceptance-criteria} i.e. integration, certification, pre--production and
production. Starting with {\sl EGEE--II}, the project adopted automation in the software
lifecycle process by leveraging the automatic build system for Grid middleware, that is
the {\sl ETICS} \cite{etics} solution.

\subsection{E--Infrastructure for Testing, Integration and Configuration of Software}

The E--Infrastructure for Testing, Integration and Configuration of Software~\cite{etics}
({\sl ETICS}, Jan 2006 -- Feb 2010) project aimed
at addressing the challenges in producing quality software in distributed,
collaborative projects such as {\sl EGEE} and its {\sl gLite} middleware. The ETICS framework
integrated different technologies and tools in order to provide automated configuration,
build and testing capabilities, as well as auto--generated documentation and
software metrics gathering -- such as Source Lines Of Code (SLOC), complexity and
number of defects/bugs \cite{etics} --. The {\sl ETICS} framework was the first automated
service for delivering quality software products in distributed environments like
the Grids.

\subsection{European Middleware Initiative}

The European Middleware Initiative ({\sl EMI}, May 2010 -- Apr 2013)
\cite{cordis:emi} project joined the 4 major Grid middleware providers in
Europe at the time -- {\sl gLite}, {\sl UNICORE}, {\sl ARC} and {\sl dCache} --
with the goal to maintain and evolve the middleware focusing on extending their interoperability and improving the reliability of the services. The
ISO/IEC 9126 \cite{iso-9126} standard was used in order to identify a set of
characteristics that needed to be present in the {\sl EMI} software products and
processes to be able to meet the {\sl EMI} quality requirements
\cite{emi-quality-model}.

For each software characteristic, a set of associated
metrics and Key Performance Indicators (KPIs) were identified and defined in
detail in the {\sl EMI} Metrics Specification \cite{emi-quality-model}. The project
leveraged the {\sl ETICS} service for the development, continuous integration and release management,
as well as for metric tracking, making queries on the collected data to display
them through a chart generation framework.

\subsection{EGI--Integrated Sustainable Pan--European Infrastructure}

The {\sl EGI--InSPIRE} (Integrated Sustainable Pan--European Infrastructure for
Researchers in Europe, May 2010 -- May 2014) project \cite{cordis:egi-inspire}
was the continuation of the {\sl EGEE--III} project, with the objective of establishing and
maintaining a sustainable European Grid Infrastructure, composed by a federation of National
Grid Initiatives and interoperable with other Grids worldwide.
This goal would be accomplished through the
development and maintenance of various operational tools -- such as the Operations Portal \cite{egi-ops},
the {\sl EGI} Helpdesk \cite{ggus}, or the Grid configuration database (GOCDB) \cite{gocdb} -- and the
management of the software provisioning process (SWPP) \cite{mario}, dealing with the validation and distribution of the software to
the production infrastructures. As a result, this process led to a production--ready
Grid computing middleware distribution named {\sl UMD} (Unified Middleware Distribution). At that time the {\sl UMD}
was a stack of about 250 software components from several technology providers, such
as {\sl EMI} and {\sl Globus}. In order to be distributed through UMD
repositories, the software had to be compliant with EGI's Quality Criteria (QC) \cite{egi-qc}. The
QC defined a set of requirements in different areas: documentation, deployment, security, information
model and operations.

The impact of the quality assurance activities was observed once the EMI software contributed to the
EGI--InSPIRE project. The data reported in Table \ref{tab:emi} (see page 13 in \cite{emi-quality-model})
summarizes the EMI software quality evolution per project quarters (PQ) as evaluated by the EGI project
with the help of the UMD QC. We can observe the improvement in quality of the EMI releases over time,
measured by the number of products that both met the software distribution criteria, by means of the UMD QC
definition, and the EGI Stage Rollout (SR) phase. This latter phase completed the criteria validation by
deploying the UMD QC--certified software in a set of candidate production sites.

\begin{table}[!h]
\renewcommand{\arraystretch}{1.3}
\caption{EMI software quality evolution}
\label{tab:emi}
\centering
\begin{tabular}{lp{2.5cm}p{2.2cm}p{1.9cm}p{2.1cm}}
\hline
\hline
PQ & Release Products (RP) & N. RP Passed UMD QC & N. RP Passed SR & N. RP Failed UMD QC\\
\hline
\hline
5 & 30 & 27 & 27 & 0\\
6 & 30 & 28 & 26 & 2 \\
7 & 27 & 26& 24& 2\\
8 & 18 & 18 & 18 & 0\\
\hline
\hline
\end{tabular}
\end{table}

{\sl UMD} is still being used and
deployed in the European scientific e--Infrastructures under the follow--up
project {\sl EGI--Engage} (Engaging the EGI Community towards an Open Science
Commons) \cite{cordis:egi-engage}. This distribution is currently complemented
by a Cloud--specific one called {\sl CMD} (Cloud Middleware Distribution). The
increase in the number of products that the new CMD distribution brought in,
compromised the effectiveness of the SWPP realization, in particular the
software validation phase. The modernization of the EGI QC validation was
accomplished through the adoption of automation, by means of the programmatic
evaluation of its fundamental quality requirements \cite{orviz2018umd}.

\subsection{INtegrating Distributed data Infrastructures for Global explOitation}

The {\sl INDIGO--DataCloud} (INtegrating Distributed data Infrastructures for Global
explOitation, Apr 2015 -- Sep 2017) \cite{salomoni2018indigo} is the last
project of software development considered in this paper. The project addressed
the existing gaps at the Cloud Platform--as--a--Service (PaaS) and Software--as--a--Service (SaaS) levels,
helping developers, e--Infrastructure administrators and scientific communities to exploit
Cloud computing benefits.

\section{New trends in software reliability}
\label{sec:ntsr}

{\sl INDIGO--DataCloud} is the most recent project included in this paper, it
leverages the lessons learned from the previous experiences described in
Section~\ref{sec:ev} and, consequently, it will be thoroughly analyzed. The expertise gathered throughout these years was highly
profitable regarding evaluation and application of innovative software engineering
methodologies, management of new technologies to put those methodologies
into practice, compilation and analysis of an appropriate set of metrics to measure the
quality of the implemented solutions.
An additional level of complexity, comes from the fact that these large collaborative projects
have an extensive list of partners.

In the following sections, we describe the three main pillars that sustain the project's strategy to attain reliability throughout the software lifecycle:

\begin{enumerate}
    \item The definition of a \textit{Software Quality Assurance (SQA) criteria and a metrics gathering procedure} to provide guidelines
    aiming at developing quality software, validating each change in the code, to facilitate its adoption. The software lifecycle is
    continuously monitored by a set of metrics, allowing prompt reactions to detected malfunctions that may occur throughout the process.
    \item The \textit{promotion of automation} to enable the foregoing change--based validation approach and to accelerate the delivery
    process of new software versions.
    \item A post-release validation to be carried out over the project's \textit{preview testbeds}, where the new product versions
    are deployed as part of the integration validation, and the \textit{testing in production environments} by external resource providers.
\end{enumerate}

\begin{figure}[ht]
\centering
\includegraphics[width=0.8\textwidth]{images/codestyle.png}
\caption{Code style standards followed by {\sl INDIGO--DataCloud}'s software products.}
\label{fig:fig_codestyle}
\end{figure}

\begin{figure*}[ht]
\centering
\includegraphics[width=\textwidth]{images/fig2.png}
\caption{Comparison of the unit testing coverage values for the {\sl INDIGO--DataCloud}
software stack over the {\sl INDIGO--1} and {\sl INDIGO--2} major releases. Products that
were exclusively released during {\sl INDIGO--2} do not show data for the first period of
{\sl INDIGO--1}.}
\label{fig:fig_unittest}
\end{figure*}

\begin{figure}[ht]
\centering
\includegraphics[width=0.8\textwidth]{images/fig3.png}
\caption{Adoption of Configuration Management tools throughout the project lifetime. The figure shows the trend lines
leading to the first (light cream points and line) and second release (dark blue points and line).}
\label{fig:fig_confman}
\end{figure}

\subsection{Software Quality Procedures}\label{software-quality-procedures}
\label{subsec:sqa}

The software quality policies were initially defined in the first deliverable document of the
project \cite{indigo-d31}. The policies were reviewed periodically taking into account the requirements
from the user communities, the feedback from the development teams and the insights of software
engineering practices. These policies covered 1) the identification
and description of the \emph{SQA requirements} that the produced software needs
to comply with, and 2) the \emph{quality metrics}, to monitor each software product's
behaviour throughout the development, release and post--release stages.

\subsubsection{SQA criteria}

The SQA criteria are a set of conventions and recommendations that pave the way for
the adequate development, timely delivery and reliable operation of the produced software components.
They emphasize the quality requirements and best practices to be applied at the
development phase, such as style and testing compliance or human code reviews, to protect the
production source code versions. Furthermore, the criteria promote the adoption of the software
by providing minimum requirements of documentation content and stressing the usage of automated
solutions for software deployment. The SQA criteria are publicly available \cite{sqa-baseline} and
it is continuously evolving by means of an open collaboration procedure \cite{sqa-baseline-repo}
that aims at serving as a reference for quality assurance realization in research software.

\begin{itemize}
\item \textit{Code style}.
The ultimate goal for code style assessment is to improve the readability and reusability of the
source code produced under the scope of the
project. Code style standards are enforced for every software component in the project stack. Community
most--adopted or \textit{de--facto standards} are initially recommended, however the development teams eventually
choose the set of guidelines to comply with. Fig.~\ref{fig:fig_codestyle} highlights the code style standards and
their popularity
in the {\sl INDIGO--DataCloud} software stack.

\item \textit{Unit and Functional testing}.
Changes involving the addition of new functionalities are required to be tested. \textit{Regression
testing} in this context is accomplished by enforcing the definition and periodic execution
of the tests, covering past fixed issues, to ensure that they are not reintroduced during the development
activities. \textit{Unit testing} completes the source code testing evaluation by focusing on the code's
internal design. The SQA criteria of {\sl INDIGO--DataCloud} set a recommended threshold of 70\% coverage
for unit testing. In Fig.~\ref{fig:fig_unittest}, unit testing coverage is compared for each software
component between the two major releases, denoting an incremental trend over the course of the project:
code coverage increased from an average value of 50.32\% to 63.34\%, outlined by the respective dates of
the project's first (INDIGO--1) and second (INDIGO--2) releases. Nevertheless, Fig.~\ref{fig:fig_unittest}
shows a decrease in the unit testing coverage values for 18\% of products. As unit tests cover low-level
code elements, they are best considered while the new code is being written \cite{unit-test-frameworks}.
We have observed that this decrease was aligned with high-demanding periods of software development, as
seen in the previous months of the INDIGO-2 major release. Regardless of the potentially higher defect
density that this fact brought along, the overall performance kept progressing as the rest of the software
stack substantially improved their own unit testing coverage statistics. By INDIGO--2, 53\% of the software
components (17 out of 32) were over the threshold of the aforementioned project's SQA code coverage
recommendation (70\%), while 75\% (24 out of 32) of the product stack exceeded 50\% coverage.
The individual records followed also a growing trend (only 6 out of 32 components had lower values in
INDIGO--2). It is worth noting that Fig.~\ref{fig:fig_unittest} not only includes new software developments
implemented from scratch within the project. Software tools and libraries from
external open--source projects -- such as {\sl tosca-parser} or {\sl heat-translator} --, contributed upstream
by the project, and well--established products, involved in {\sl INDIGO--DataCloud}
but not contributing with the 100\% of their codebases -- such as {\sl dCache} --, are also considered in the
analysis. However the values given apply to the entire codebase, even for the latter case of products. These
products challenged the application of the SQA policies described in this section, by two means:
\begin{itemize}
    \item Previously developed code was not refactored since it was not \textit{owned} by the project.
    \item An agreement on prevailing SQA policies for such cases where quality practices were already in place.
\end{itemize}

\item \textit{Integration testing}. Software components usually interact with other services during
operation. Integration testing deals with the interactions among coupled software components or
parts of a system that cooperate to achieve a given functionality. This type of testing might be
complex and, based on the project's experience, difficult to be implemented in an automatic way. The aim is to
guarantee the overall operation of the component with regard to the services it interfaces with,
whenever new functionalities are introduced.

\item \textit{Code review}.
This phase is the last step in the change management pipeline, once the candidate change has
successfully passed through the testing methods described previously. It implies the human--based
revision of the proposed change to discuss its adequacy in terms of e.g. scope, objective fulfilment,
and documentation completeness. On approval, the candidate change is definitely merged into
the source code's production version. Human code reviewing is paramount in the software quality assessment, specially
important when the SQA testing requirements are fully automated. Secure code reviews are also
done at this stage, assessing common vulnerabilities from inputs coming from automated linters
and manual dynamic application security testing.

\item \textit{Documentation}.
The SQA criteria set the path for the adoption of the developed software by
defining the documentation content, according to the target audience. This requirement also
promotes the automation, both in terms of documentation creation and service deployment. The
\textit{documentation is treated as code}, using a markup language, automatically rendered and
uploaded to online repositories \cite{foot1}.
Thus documentation is portable and
human--reviewed, using the same workflow as the code does, validating the changes before being
updated into the production repository.

\item \textit{Automated Deployment}.
To lower the barriers of software adoption, the SQA criteria require the automated
deployment of the products delivered as part of the catalogue. Automation in this context is
tackled using configuration management tools. These tools allow the adoption of
\textit{Infrastructure as Code} (IaC) practice, managing the component's deployment through declarative
definitions. The definition files appear as an additional source of documentation
-- self--documenting code -- as they sequentially guide the component's deployment process on multiple
platforms. {\sl INDIGO--DataCloud} project contributed to open--source IaC tools such
as Ansible
and Puppet \cite{foot2}. A representative example of
such contributions are the \textit{50 roles} developed from scratch and currently hosted in
the Ansible Galaxy portal \cite{foot3}. Fig.~\ref{fig:fig_confman}
shows the number of products that offer an automated means for deployment.
It shows an increase in the adoption of such tools leading to the {\sl INDIGO--1} release
(light cream points and trend line), as well as to the {\sl INDIGO--2} release
(dark blue points and trend line). The rate of adoption is lower during the weeks
before the second release because a significant fraction of the products had already adopted it
previously to the first release.
\end{itemize}

\subsubsection{Quality metrics}

The evaluation of the software quality is performed by measuring the values of
the metrics and Key Performance Indicators (KPIs) defined based upon the
ISO/IEC 9126 standard. These metrics cover the development, release and
maintenance phases of the software lifecycle.

\textit{Development metrics} are obtained programmatically from several sources, namely GitHub
API \cite{foot4} and the Jenkins \cite{foot5} service, and graphically displayed as GitHub
pages using the GrimoireLab framework \cite{foot6}.
Per--component weekly reports, including the SQA requirement
fulfilment, are issued and individually discussed through the different communication channels
with the development teams. Issue tracking metrics and KPIs are an essential piece of information
of both feature addition and defect solving, useful to detect and fix misbehaviours while in the
development phase.

\textit{Release metric} sources are the online repository servers, namely the
Linux package \cite{foot7} and DockerHub \cite{foot8} repositories. Jenkins
server also contains valuable release data as it is the service where the packages
and containers are being built before being uploaded to the online repositories. Release
KPIs primarily focus on the frequency and efficiency -- mainly rollbacks -- statistics.

\textit{Maintenance and user support metrics} are key for continuously improving the response
to issues reported by external users. Feedback is collected from outside helpdesks, such as
EGI's GGUS~\cite{ggus}, and the GitHub Issues tracker.

\subsection{Software quality validation and delivery automation}\label{software-automation}
\label{sec:devops}
Tackling the granularity of the former SQA requirements -- most of which accounted in a per--change basis --
is hardly achieved without the aid of automation. The introduction of automation to validate the quality
requirements increases the overall reliability of the produced software: automated testing is more
time-efficient, leading to higher code coverages and increased defect detection, when compared with a manual
approach \cite{rafi2012}. To put it in numbers, as Fig.~\ref{fig:indigobugs} showcases, the
total number of defects detected in the pre--production stages surpassed by a factor of 30
the ones reported by external users, over the lifetime of the {\sl INDIGO-DataCloud}
project.
\begin{figure}[h]
\centering
\includegraphics[width=0.8\textwidth, height=50mm]{images/indigoBugs.png}
\caption{Number of software bugs -- documentation typos are excluded -- detected in development 
and production stages over the lifetime of the {\sl INDIGO-DataCloud} project. Production 
defects correspond to software bugs filed by users of the EGI e--Infrastructure throughout 
the next 2 years after the {\sl INDIGO--1} major release, in August 2016 (source: EGI's GGUS
helpdesk). The development defects stem only from the pre--release testing phase, filed 
exclusively as part of the SQA team work (source: GitHub Issues tracker). As a consequence, the 
development defects started to arise in the previous months prior to the {\sl INDIGO--1} release, 
while the major activity was concentrated in the months that preceded the second major release 
of {\sl INDIGO--2}, in April 2017. As it can be observed, the offset between the advent of defects
in development and production corresponds to nearly half a year, with the first production defect
reported 3 months after the {\sl INDIGO--1} release.}

\label{fig:indigobugs}
\end{figure}

As any new change, involving a new feature or fix, passes through the automated
SQA machinery, the software is always ready to be released. As a result, the often compromised harmonization
between the continuous release of new features -- pushed by the development teams -- and the maintenance of
stable production systems -- demanded by users and resource providers -- can be guaranteed. The DevOps culture
theorizes and provides practical solutions that aim at unifying both software development (dev) and software
operations (ops) teams, by emphasizing the SQA techniques to avoid infrastructure disruption whenever new
developments are deployed into production systems.
The incremental adoption of DevOps approaches throughout the project's lifetime was a key player towards
software reliability. Starting with a Continuous Integration (CI) scenario at the first stages of the project, the software delivered
was eventually validated and distributed using Continuous Delivery (CD) pipelines.

\begin{figure}[ht]
\centering
\includegraphics[width=0.8\textwidth, height=50mm]{images/jenkins_CI_builds.png}
\caption{Evolution of the total number of testing builds triggered automatically as part of the Jenkins CI implementation.}
\label{fig:fig_jenkins_CI_builds}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=\textwidth]{images/devops.png}
\caption{Continuous Delivery workflow for Docker images.}
\label{fig:fig_CD}
\end{figure}

\subsubsection{Continuous Integration (CI)}
\label{subsec:ci}

The {\sl INDIGO--DataCloud} project promoted the application of a CI scenario that enforced,
for any piece of produced software, the testing requirements defined in the SQA criteria,
as described in Section \ref{subsec:sqa}. Such an environment requires an automation
ready--to--go infrastructure where the different services and technologies involved interact with each
other to trigger the source code validation pipeline for each candidate change. The
pipeline is mainly comprised by the code style validation, unit and functional testing
coverage. The CI pipeline is complemented with additional quality checks, such as
integration tests for the software components where automation of this type of testing
is applicable, or security linters for the static code analysis of suspicious
constructs that could lead to security risks. Metrics gathering is also a part of
the CI pipeline to have a per--change trend evolution of the common source code
related metrics, such as SLOC or the cyclomatic complexity.

The practical implementation of the CI scenario leverage from tightly integrated open
source tools such as:
\begin{itemize}
    \item GitHub \cite{foot9} as the online source code repository hosting service,
    \item Jenkins as the event--response CI,
    \item Docker container provisioning system, to provide instant computing power needed
          for executing the pipeline jobs.
\end{itemize}
The project defined a source code contribution workflow, based on GitHub Pull Requests
(PRs), where each change automatically triggers in Jenkins the associated quality and
metric tracking checks on PR creation or update.

The CI pipeline is composed from a set of required checks, their exit status
is reported back to GitHub. The changes can be prevented if those tests are
not successfully executed.
As each change is validated, the chances
of early detection of defects increase. Within this scenario, the cost of defect solving
is dramatically reduced and the reliability of the software solutions improved, as any
bug or design issue is likely to be detected and subsequently corrected in this phase.

Fig.~\ref{fig:fig_jenkins_CI_builds} shows the evolution, throughout the
{\sl INDIGO--DataCloud}'s first and second releases, of the required test types builds since the
implementation of the CI infrastructure. Automated functional testing coverage were not
available for all the software stack, thus the associated number of builds are fewer.
Towards the end of the project, the software development activity slowed the pace but not completely
ceased. The support of the CI infrastructure, once reached the project's end of life, allowed the
continued maintenance and development of the prevailing software. Some development teams deployed
parallel CI systems, taking advantage of the experience gained during the project.

\begin{figure*}[ht]
	\centering
	\begin{subfigure}
		\centering
\includegraphics[width=0.85\textwidth]{images/disvis-flow.png}
\caption{DevOps pipeline to distribute Docker images for Disvis application.}
\label{fig:fig_disvis}
	\end{subfigure}
	\quad
	\begin{subfigure}
		\centering
\includegraphics[width=\textwidth]{images/pilotpreview.png}
\caption{Resource centers supporting the Pilot Preview testbed and corresponding
set of deployed {\sl INDIGO--DataCloud} components or services.}
\label{fig:fig_pilotpreview}
	\end{subfigure}
\end{figure*}

\subsubsection{Continuous Delivery}
As DevOps suggests, frequent releases positively affect the reliability of the software as
they reduce the time in which a given defect is exposed or has an impact, allowing the
development teams to act promptly based on the regular feedback \cite{chen2015}. The software updates of
{\sl INDIGO--DataCloud} products, which have been taking place since the second major release, are passing through
a Continuous Delivery (CD) pipeline that adds the packaging of the software right after the
successful execution of the CI pipeline previously described. Consequently, CI--validated
software components are automatically delivered in online repositories. Nonetheless,
expert supervision to validate the results is required.

{\sl INDIGO--DataCloud} software stack is delivered in the form of Linux software packages (rpms, debs)
and Docker containers. The CD approach is different for each type of software packaging. \textit{Linux
package pipeline} intentionally delivers the packages produced in a pre--production -- preview --
branch. Before that, once the software packages have been created, they are uploaded to a
testing branch. The available automated deployment solution -- see Section \ref{subsec:sqa} -- checks
the installation and configuration of the component relying on the testing repository. On
successful completion, the packages are automatically signed and subsequently moved to
the preview branch.
In the last step, the release manager supervises the process to move the  packages to the
production branch.

The \textit{Docker container pipeline} uses the automated deployment solution to
actually install and configure the software component in the container image. Once this is
done the recently created image is uploaded to the production DockerHub repository, tagged
with the label corresponding to the current {\sl INDIGO--DataCloud} release.
Fig.~\ref{fig:fig_CD} shows the complete workflow of the Docker container CD implementation,
automatically triggered by a change in the source code.

\subsubsection{DevOps adoption from user communities}

The experience gathered throughout the project regarding the adoption of
different DevOps practices is not only useful and suitable for the software related
to the core services in the {\sl INDIGO--DataCloud} solution, but also applicable to the
development and distribution of the applications coming from the user communities.

Two applications coming from the supported research communities,
DisVis \cite{disvis} and PowerFit \cite{powerfit}, were
integrated into a similar CI/CD pipeline described in section \ref{sec:devops}.
Fig.~\ref{fig:fig_disvis} shows the pipeline for the DisVis application.

User application developers were provided with both a means to validate the
source code before merging and the creation of a new versioned Docker image,
automatically available in the {\sl INDIGO--DataCloud}'s application repository.

The novelty introduced in the pipeline above is the validation of the application.
Once the application is packaged as a Docker image, and subsequently uploaded
to the DockerHub repository, it is instantiated in a new container to be validated.
The application is then executed and the results compared with a set of reference outputs.
This pipeline implementation goes a step forward by testing the application
execution for the latest available Docker image in the catalogue.

\subsection{Integration, preview and early adoption}

Two pilot infrastructures are at the disposal of developers and scientific
communities involved in the project. The aim of these testbeds is to test the
level of integration between the components involved in the {\sl INDIGO--DataCloud}
solutions and use cases validation. Within these testbeds, the applications are deployed
and executed
with the last stable version of the software in environments as similar as
possible to production ones. A map of the Pilot Preview
infrastructure is depicted in Fig.~\ref{fig:fig_pilotpreview}. It shows the
resource providers and the components or services they deployed and supported.

The released software is tested in production environments through the
Staged--Rollout process, before its inclusion in the EGI's Cloud Middleware Distribution (CMD).
Selected resource providers are requested to install
the most updated stable versions of the software, and to give access to them to their users. The
Staged--Rollout process is key to detect and mitigate issues that could only
appear in production environments.

\section{Conclusion}
\label{sec:con}

%After more than a decade of experiences collected throughout the reviewed
%EU--funded software development and e--Infrastructure management projects, it
%becomes apparent that the European software produced for scientific purposes is
%evolving to a more sustainable model where the quality and reliability of
%software is being prioritized. The gradual adoption of software engineering insights, such as
%DevOps practices or Agile methodologies, are leading the change. In this regard, the continuous evolution of the open--source
%collaborative tools, extending their capabilities to bring along tight
%integrations amongst other services, matched the requirements of such practices
%and procedures, facilitating in a great manner their implementation in real scenarios.

% Comment
%%P8 C1 L26: What is the demonstration that reliability has been improved significantly with the addition of stricter methologies? Is there some quantitative evidence of uptime, or bug reports per user, or similar? Or some evidence of speed of defect detection & fixing, which you cite as greatly improved? I am very sympathetic to the idea that systematic development techniques can help improve software quality (without getting into circular definitions, where "quality" means that methodologies or standards have been applied!) but there will also be a point at which the methodology gets too heavy and compromises the "agility" of the development team: how can you identify which processes are helping and which are not?
Both reliability and sustainability of the {\sl INDIGO--DataCloud} software products 
have been primarily improved by acting at the source code level. According to the 
size of the project, which comprised more than 30 products and 200 GitHub repositories, 
this strategy required a remarkable preliminary ground work. On the one hand, the 
definition of a set of SQA requirements -- compiled in the described SQA criteria 
-- that shall be fulfilled prior to release by the {\sl INDIGO--DataCloud} core 
products. On the other hand, the need of setting up the testing infrastructure and
the associated CI/CD pipelines to check that the changes in the source code matched 
those requirements in an automated fashion. As a complement to the above strategy, a 
final validation step combined the deployment of the pre--released software versions 
in the project's preview testbed, and subsequently, most of the products were tested
in selected production environments through the aid of EGI's staged--rollout process.

The most noticeable metric that demonstrates a reliability improvement comes from the 
fact that the number of software defects uncovered throughout the pre--release phase 
surpassed by a factor of 30 the ones reported by end users in the EGI e-Infrastructure
during the next two years after the first major release (INDIGO-1). Accordingly, the
sustainability of the software was ...

Nevertheless, the individual products of {\sl INDIGO--DataCloud} did not undergo 
a progressive quality improvement throughout the lifespan of the project. Indeed,
unit testing coverage dropped for about 18\% of the total products halfway through
the project. This behavior was aligned with the high-demanding periods of software
development, in particular, the previous months before the second major release 
(INDIGO--2). This fact demonstrates that, even considered as best practice, test 
cases are not commonly written right after, or even before, introducing a new unit
or functionality in the code. High coverage values are generally advisable but they
entail time--consuming work that is often neglected by the code owners. The effort 
spent in testing should be oriented to those sections of the code that are more 
likely to cause negative effects to end users. Consequently, a high coverage value 
does not generally provide a linear benefit with respect to the effort spent -- the
SQA criteria recommends 70\% of coverage as a good compromise -- and, in addition,
it does not indicate whether the key sections of the code are covered. However, unit
testing is a valuable tool to detect defects at coding time, so whenever the test 
cases accompany the changes done in the code, a high unit testing coverage value is 
a good indicator of reliability in the software.
%(related defect increase?, resulting from the INDIGO-2 release?)

% products not developed from scratch challenged the application of SQA policies??

%

applying the SQA criteria in the development stage. As each change in 
the source code is validated to match the SQA requirements, defects in the software 
are more likely to be detected and fixed at the earliest stage. Consequently, costs
are reduced, since solving bugs at later stages require greater efforts, and the end
user obtains a much more reliable piece of software, as demonstrated by the low rate
of defects found once it is in production. Automation is of paramount importance to 
guarantee the {\sl INDIGO--DataCloud}'s SQA criteria compliance. Thus, through the
application of DevOps practices, a CI/CD scenario was put in place to cover both the
validation and delivery of the {\sl INDIGO--DataCloud} software products. The CI/CD 
pipelines provide the agility to produce releases frequently enough to reduce the 
impact on production systems, since the software gets progressively updated.

In this regard, a pre--production validation,
through the preview testbeds or resource center's early adoption -- handled by the
Staged--rollout process -- alleviates the likelihood of discovering uncovered
issues at production.

The software lifecycle demands as well a continuous measurement and analysis
process, which allows timely reactions to discover and fix defects throughout the
development, release and maintenance stages. Conscientious metric
tracking and response sits at the core of an appropriate software maturity, as
ultimate reliability can only be achieved by enforcing and empowering the SQA
criteria through each stage of the software lifecycle.

Lessons learned show that sustainability is a key pursuit. The achievements, in terms of
processes, knowledge, software and infrastructures, shall be profited from former successful
completed projects. In this regard, the more recent project described in this paper, {\sl INDIGO-DataCloud},
may have set the path for future reference in the European software development research ecosystem.
The SQA criteria, describing the quality requirements that might be considered during the entire
software lifecycle, have been published as open access and periodically reviewed to incorporate
new recommendations and best practices. Successful software products created during the lifetime of the
project are maintained leveraging the same source code repositories and CI infrastructure. A seamless
knowledge transfer among subsequent software development initiatives is key for the good progress of
delivering reliable software.

\section*{Acknowledgment}

{\sl DataGrid - Research and Technological Development for an International Data Grid}
project has received funding from the European Union's Fifth Framework Programme under
grant agreement IST--2000--25182E.

{\sl EGEE - Enabling grids For E--science} project has received funding from the European
Union's Sixth Framework Programme under grant agreement INFSO--RI--508833.

{\sl EGEE - Enabling grids For E--science--II} project has received funding from the
European Union's Sixth Framework Programme under grant agreement INFSO--RI--031688.

{\sl EGEE - Enabling grids For E--science--III} project has received funding from the
European Union's Seventh Framework Programme under grant agreement INFSO--RI--222667.

{\sl ETICS - E--Infrastructure for Testing, Integration and Configuration of Software}
project has received funding from the European Union's Sixth Framework Programme under
grant agreement INFSO--RI--026753.

{\sl ETICS - E--Infrastructure for Testing, Integration and Configuration of Software - Phase 2}
project has received funding from the European Union's Seventh Framework Programme under grant
agreement INFSO--RI--223782.

{\sl EGI--InSPIRE - European Grid Initiative: Integrated Sustainable Pan--European
Infrastructure for Researchers in Europe} project has received funding from the European
Union's Seventh Framework Programme under grant agreement INFSO--RI--261323.

{\sl EGI--Engage} project has received funding from the European Union's Horizon 2020
research and innovation programme under grant agreement RIA--654142.

{\sl INDIGO--DataCloud} project has received funding from the European Union's Horizon
2020 research and innovation programme under grant agreement RIA--653549.

\begin{thebibliography}{1}

\bibitem{h2020}
\emph{Horizon 2020} -- European Commission.
[Online]. Available: \url{https://ec.europa.eu/programmes/horizon2020/}.
[Accessed 14 Feb. 2019].

\bibitem{lingrand-egee}
D. Lingrand, J. Montagnat, J. Martyniak and D. Colling,
``Analyzing the EGEE production grid workload: application to jobs submission optimization,''
\emph{Workshop on Job Scheduling Strategies for Parallel Processing},
pp. 37--58, May. 2009.

\bibitem{campana-atlas}
S. Campana  \emph{et al.},
``Analysis of the ATLAS Rome production experience on the LHC computing grid,''
\emph{IEEE 1st Int. Conf. of e-Science and Grid Computing},
pp. 8-pp, 2005.

\bibitem{kindermann}
S. Kindermann,
``Climate Data Analysis and Grid Infrastructures: Experiences and Perspectives,''
\emph{Grid-Enabling Legacy Applications and Supporting End Users Workshop (GELA)},
vol. 20. 2006.

\bibitem{mendez-lorenzo}
P. Mendez-Lorenzo, J. T. Moscicki and A. Ribon,
``Experiences in the gridification of the Geant4 toolkit in the WLCG/EGEE environment,''
\emph{IEEE Nucl. Sci. Symp. Conf. Rec.},
vol. 2, 2006.

\bibitem{aiftimiei}
C Aiftimiei, A Ceccanti, D Dongiovanni, A Di Meglio and F Giacomini,
``Improving the quality of EMI Releases by leveraging the EMI Testing Infrastructure,''
\emph{J. Phys. Conf. Ser.},
vol. 396, no. 5, 2012.

\bibitem{agile-manifesto}
K. Beck \emph{et al.},
``Manifesto for Agile Software Development,'' 2012.
[Online]. Available: \url{http://www.agilemanifesto.org/}.
[Accessed 14 Feb. 2019].

\bibitem{zhu}
L. Zhu, L. Bass and G. Champlin--Scharff,
``DevOps and Its Practices,''
\emph{IEEE Softw.},
vol. 33, n. 3, pp. 32--34, 2016.

\bibitem{soltesz}
S. Soltesz \emph{et al.},
``Container-based Operating System Virtualization: A Scalable, High-performance Alternative to Hypervisors,''
\emph{SIGOPS Oper. Syst. Rev.},
vol. 41, no. 3, pp. 275--287, Mar. 2007.

\bibitem{radatz}
``IEEE Standard Glossary of Software Engineering Terminology,''
\emph{IEEE Std. 610.12--1990}, pp.1--84, Dec. 31 1990.

\bibitem{sommerville}
I. Sommerville,
``Integrated requirements engineering: A tutorial,''
\emph{IEEE Softw.}, vol. 22, no. 1, pp. 16--23, Jan.--Feb. 2005.

\bibitem{tashtoush}
Y. Tashtoush, Z. Odat, I. Alsmadi, M. Yatim,
``Impact of programming features on code readability,''
\emph{Int. J. Softw. Eng. Appl.},
vol. 7, no. 6, pp. 441--458, Nov. 2013.

\bibitem{buse}
R. P. Buse and W. R. Weimer,
``Learning a metric for code readability,''
\emph{IEEE Trans. Softw. Eng.},
vol. 36, no. 4, pp. 546--558, Jul. 2010.

\bibitem{baggen}
R. Baggen, J. P. Correia, K. Schill and J. Visser,
``Standardized code quality benchmarking for improving software maintainability,''
\emph{Softw. Qual. J.},
vol. 20, no. 2, pp. 287--307, Jun. 2012.

\bibitem{fenton}
N. Fenton  and  S.  L.  Pfleeger,
``Software  Metrics:  A  Rigorous  and  Practical Approach,''
\emph{International Thomson Computer Press},
London, UK, second edition, 1997.

\bibitem{capers}
J. Capers,
``Applied Software Measurement (2nd Ed.): Assuring Productivity and Quality,''
\emph{McGraw Hill}, New York, NY, USA, 1997.

\bibitem{goodman}
P. Goodman,
``Practical Implementation of Software Metrics,''
\emph{McGraw Hill}, New York, NY, USA, 1993.

\bibitem{mccabe}
T. J. McCabe,
``A Complexity Measure,''
\emph{IEEE Trans. Softw. Eng.},
vol. 2, no. 4, pp. 308--320, Dec. 1976.

\bibitem{chidamber}
S. R. Chidamber and C. E. Kemerer,
``A metrics suite for Object--oriented design,''
\emph{IEEE Trans. Softw. Eng.},
vol. 20, no. 6, pp. 476--493, Jun 1994.

\bibitem{lorenz}
M. Lorenz and J. Kidd,
\emph{Object--Oriented Software Metrics: A Practical Approach}.
Prentice-Hall, Inc., Upper Saddle River, NJ, USA, 1994.

\bibitem{myers}
G. J. Myers, C. Sandler and T. Badgett,
\emph{The art of software testing, 3rd Ed.}.
Wiley Publishing, Dec. 2011.

\bibitem{athanasiou}
D. Athanasiou, A. Nugroho, J. Visser and A. Zaidman,
``Test Code Quality and Its Relation to Issue Handling Performance,''
\emph{IEEE Trans. Softw. Eng.},
vol. 40, no. 11, pp. 1100--1125, Nov. 2014.

\bibitem{horgan}
J. R. Horgan, S. London and M. R. Lyu,
``Achieving software quality with testing coverage measuring,''
\emph{Comput.},
vol. 27, no. 9, pp. 60--69, Sep. 1994.

\bibitem{cordis:datagrid}
P. Kunszt,
``European DataGrid project: status and plans,''
\emph{Nucl. Instr. Meth. Phys. Res. A},
vol. 502, no. 2, pp. 376--381, Apr. 2003.

\bibitem{gagliardi}
F. Gagliardi, B. Jones, M. Reale and S. Burke,
``European DataGrid Project: Experiences of Deploying a Large Scale Testbed for E--science Applications,''
in \emph{Performance Evaluation of Complex Systems: Techniques and Tools,}
Performance 2002. LNCS, vol. 2459, pp. 480--499, 2002.

\bibitem{globus}
I. Foster and C. Kesselman,
``Globus: a Metacomputing Infrastructure Toolkit,''
\emph{Int. J. High Perfor. Comput. Appl.,}
vol. 11, no. 2, pp. 115--128, Jun. 1997.

\bibitem{datagrid}
L. Momtahan and A. Martin,
``e--Science Experiences: Software Engineering Practice and the EU DataGrid,''
in \emph{Proc. 9th Asia--Pacific Softw. Eng. Conf.},
pp. 269--275, Dec. 2002.

\bibitem{agile}
T. Dingsoyr, S. Nerur, V. Balijepally and N. B. Moe,
``A decade of agile methodologies: Towards explaining agile software development,''
\emph{J. Syst. Softw.},
vol. 85, no. 6, pp. 1213--1221, Jun. 2012.

\bibitem{cmm}
M. Paulk, B. Curtis, M. Chrissis and V. C. Weber,
``Capability maturity model for software,''
Softw. Eng. Inst.,
Technical Report CMU/SEI-93-TR-024, ESC-TR-93-177, Feb. 1993.
[Online]. Available: \url{https://resources.sei.cmu.edu/asset_files/TechnicalReport/1993_005_001_16211.pdf}.
[Accessed 14 Feb. 2019].

\bibitem{datagridguide}
Quality Assurance Group, ``DataGrid - European DataGrid Developers' Guide,''
2003.
[Online] Available: \url{https://edms.cern.ch/ui/file/358824/1.1/EDG-DevGuide-v1-2.pdf}
[Accessed 14 Feb. 2019].

\bibitem{datagridindicators}
DataGrid, ``DataGrid Internal Document - Quality and Performance Indicators for DataGrid,''
2003.
[Online] Available: \url{https://edms.cern.ch/ui/file/386039/2/QIv0-3.pdf}
[Accessed 14 Feb. 2019].

\bibitem{cordis:egee}
\emph{Enabling Grids for E--sciencE (EGEE)} project, European Community Research and
Development Information Service (CORDIS).
[Online] Available: \url{http://cordis.europa.eu/project/rcn/80149_en.html}
[Accessed 14 Feb. 2019].

\bibitem{cordis:egee2}
\emph{Enabling Grids for E--sciencE--II (EGEE--II)} project, European Community Research and
Development Information Service (CORDIS).
[Online] Available: \url{http://cordis.europa.eu/project/rcn/99189_en.html}
[Accessed 14 Feb. 2019].

\bibitem{cordis:egee3}
\emph{Enabling Grids for E--sciencE--III (EGEE--III)} project, European Community
Research and Development Information Service (CORDIS).
[Online] Available: \url{http://cordis.europa.eu/project/rcn/87264_en.html}
[Accessed 14 Feb. 2019].

\bibitem{egee}
F. Gagliardi and M. E. Begin,
``EGEE -- providing a production quality grid for e-science,''
in \emph{2005 IEEE Inter. Symp. Mass Storage Syst. Technol.},
pp. 88--92, Jun. 2005.

\bibitem{glite}
E. Laure \emph{et al.},
``Programming the Grid with gLite,''
\emph{Computational Meth. Sci. Technol.},
vol. 12(1), pp. 33--45, 2006.

\bibitem{condor}
D. Thain, T. Tannenbaum and M. Livny,
``Condor and the Grid,''
in \emph{Grid Computing: Making the Global Infrastructure a Reality},
ch. 11, pp. 63--70, 2003.

\bibitem{egee:acceptance-criteria}
\emph{Definition and Documentation of the Revised Software Life--Cycle Process},
Milestone MSA3.4.2, 2010, EGEE--III project.
[Online]. Available: \url{https://edms.cern.ch/ui/file/1062487/2/EGEE-III-MSA3.4.2-1062487-v1_4.pdf}.
[Accessed 14 Feb. 2019].

\bibitem{etics}
A D Meglio, M-E Begin, P Couvares, E Ronchieri and E Takacs,
``ETICS: the international software engineering service for the grid,''
in \emph{J. Phys.: Conf. Ser.},
vol. 119, no. 4, 042010, 2008.

\bibitem{cordis:emi}
C. Aiftimiei \emph{et al.},
``Towards next generations of software for distributed infrastructures: The European Middleware Initiative,''
in \emph{2012 IEEE 8th Inter. Conf. on E-Science},
Chicago, IL, pp. 1--10, 2012.

\bibitem{iso-9126}
\emph{ISO/IEC 9126 Software Engineering - Product Quality},
International Organization for Standardization.
[Online]. Available: \url{https://www.iso.org/standard/22749.html}.
[Accessed 14 Feb. 2019].

\bibitem{emi-quality-model}
M. Alandes \emph{et al.},
``Experiences with Software Quality Metrics in the EMI middleware,''
in \emph{J. Phys.: Conf. Ser.},
vol. 396, no. 5, 052003, 2012.

\bibitem{cordis:egi-inspire}
I. C. Plasencia,
``EGI.eu the European grid initiative,''
in \emph{Proc. 4th Iberian Grid Infra. Conf.},
pp. 5--15, 2010.

\bibitem{egi-ops}
H. Cordier \emph{et al.},
``From EGEE Operations Portal towards EGI Operations Portal,''
in \emph{Data Driven e-Science (ISGC2010)},
pp. 129--140, 2011.

\bibitem{ggus}
T. Antoni, et al.,
Global grid user support--building a worldwide distributed user support infrastructure,
\emph{J. Phys.: Conf. Ser.},
vol. 119, no. 5, 052002, 2008.

\bibitem{gocdb}
G. Mathieu and J. Casson,
``GOCDB4, a New Architecture for the European Grid Infrastructure,''
in \emph{Data Driven e-Science (ISGC2010)},
pp. 163--174, 2011.

\bibitem{mario}
M. David \emph{et al.},
``Validation of Grid Middleware for the European Grid Infrastructure,''
\emph{J. Grid Comp.},
vol. 12, no. 3, pp. 543--558, 2014.

\bibitem{egi-qc}
\emph{EGI Quality Criteria}.
[Online]. Available: \url{https://egi-qc.github.io/}.
[Accessed 14 Feb. 2019].

\bibitem{cordis:egi-engage}
\emph{Engaging the EGI Community towards an Open Science Commons (EGI--ENGAGE)}
project, European Community Research and Development Information Service (CORDIS).
[Online] Available: \url{http://cordis.europa.eu/project/rcn/194937_en.html}
[Accessed 14 Feb. 2019].

\bibitem{orviz2018umd}
P. Orviz \emph{et al.},
``umd-verification: Automation of Software Validation for the EGI Federated e-Infrastructure,''
\emph{J. Grid Comp.},
vol. 16, no. 4, pp. 683--696, 2018.

\bibitem{salomoni2018indigo}
D. Salomoni \emph{et al.},
``Indigo-datacloud: a platform to facilitate seamless access to e-infrastructures,''
\emph{J. Grid Comp.},
vol. 16, no. 3, pp. 381--408, 2018.

\bibitem{indigo-d31}
J. Gomes \emph{et al.},
``Initial Plan for WP3,''
\emph{INDIGO--DataCloud Deliverable 3.1}.
[Online]. Available: \url{https://www.indigo-datacloud.eu/documents/initial-plan-wp3-d31}.
[Accessed 14 Feb. 2019].

\bibitem{sqa-baseline}
Pablo Orviz \emph{et al.},
``A set of common software quality assurance baseline criteria for research projects,''
2017.
[Online]. Available: \url{http://hdl.handle.net/10261/160086}.
[Accessed 14 Feb. 2019].

\bibitem{sqa-baseline-repo}
Members of the INDIGO-DataCloud, DEEP Hybrid-DataCloud and eXtreme DataCloud collaborations, 2015-2020,
``A set of Common Software Quality Assurance Baseline Criteria for Research Projects,''
2018.
[Online]. Available: \url{https://github.com/indigo-dc/sqa-baseline}
[Accessed 14 Feb. 2019].

\bibitem{unit-test-frameworks}
Paul Hamill,
``Unit test frameworks: tools for high-quality software development,''
\emph{O'Reilly Media, Inc.},
2004.

\bibitem{foot1}
indigo-dc, ``indigo-dc Spaces - GitBook,''
2018.
[Online]. Available: \url{https://www.gitbook.com/@indigo-dc}.
[Accessed 14 Feb. 2019].

\bibitem{foot2}
puppetforge, ``Modules by INDIGO Datacloud - Puppet Forge,''
2018.
[Online]. Available: \url{https://forge.puppet.com/indigodc}.
[Accessed 14 Feb. 2019].

\bibitem{foot3}
GALAXY, ``Ansible Galaxy,''
2018.
[Online]. Available: \url{https://galaxy.ansible.com/indigo-dc/}
[Accessed 14 Feb. 2019].

\bibitem{foot4}
GitHub Developer, ``GitHub API v3 | GitHub Developer Guide,''
2018.
[Online]. Available: \url{https://developer.github.com/v3/}
[Accessed 14 Feb. 2019].

\bibitem{foot5}
Jenkins, ``Jenkins,''
2018.
[Online]. Available: \url{https://jenkins.io/}
[Accessed 14 Feb. 2019].

\bibitem{foot6}
GRIMOIRELAB, ``GrimoireLab - Software Development and Community Analytics Platform,''
2017.
[Online]. Available: \url{http://grimoirelab.github.io/}
[Accessed 14 Feb. 2019].

\bibitem{foot7}
Jenkins Indigo-dc, ``Jenkins - Indigo-DataCloud,''
2018.
[Online]. Available: \url{https://jenkins.indigo-datacloud.eu:8080/}
[Accessed 14 Feb. 2019].

\bibitem{foot8}
Indigo-dc, ``indigodatacloud - Docker Hub,''
2018.
[Online]. Available: \url{https://hub.docker.com/u/indigodatacloud}
[Accessed 14 Feb. 2019].

\bibitem{rafi2012}
D. M. Rafi \emph{el al.},
``Benefits and limitations of automated software testing: Systematic literature review and practitioner survey,''
in \emph{Proc. 7th Int. Workshop Automation Softw. Test},
36--42, 2012.

\bibitem{foot9}
GitHub, ``Learn Git and GitHub without any code!,''
2018.
[Online]. Available: \url{https://github.com/}
[Accessed 14 Feb. 2019].

\bibitem{chen2015}
L. Chen,
``Continuous delivery: Huge benefits, but challenges too,''
\emph{IEEE Softw.},
vol. 32, no 2, pp. 50--54, 2015.

\bibitem{disvis}
G.C.P. Van Zundert and A.M.J.J. Bonvin,
``DisVis: quantifying and visualizing the accessible interaction space of distance
restrained biomolecular complexes,''
\emph{Bioinformatics}, vol. 31, no. 19, pp. 3222--3224, 2015.

\bibitem{powerfit}
G.C.P. Van Zundert and A.M.J.J. Bonvin,
``Fast and sensitive rigid--body fitting into cryo--EM density maps with PowerFit,''
\emph{AIMS Biophys.}, vol. 2, no. 20150273, 73--87, 2015.

\end{thebibliography}

%
%
%\section{Section title}
%\label{sec:1}
%Text with citations \cite{RefB} and \cite{RefJ}.
%\subsection{Subsection title}
%\label{sec:2}
%as required. Don't forget to give each section
%and subsection a unique label (see Sect.~\ref{sec:1}).
%\paragraph{Paragraph headings} Use paragraph headings as needed.
%\begin{equation}
%a^2+b^2=c^2
%\end{equation}
%
%% For one-column wide figures use
%\begin{figure}
%% Use the relevant command to insert your figure file.
%% For example, with the graphicx package use
%  \includegraphics{example.eps}
%% figure caption is below the figure
%\caption{Please write your figure caption here}
%\label{fig:1}       % Give a unique label
%\end{figure}
%%
%% For two-column wide figures use
%\begin{figure*}
%% Use the relevant command to insert your figure file.
%% For example, with the graphicx package use
%  \includegraphics[width=0.75\textwidth]{example.eps}
%% figure caption is below the figure
%\caption{Please write your figure caption here}
%\label{fig:2}       % Give a unique label
%\end{figure*}
%%
%% For tables use
%\begin{table}
%% table caption is above the table
%\caption{Please write your table caption here}
%\label{tab:1}       % Give a unique label
%% For LaTeX tables use
%\begin{tabular}{lll}
%\hline\noalign{\smallskip}
%first & second & third  \\
%\noalign{\smallskip}\hline\noalign{\smallskip}
%number & number & number \\
%number & number & number \\
%\noalign{\smallskip}\hline
%\end{tabular}
%\end{table}

%\begin{acknowledgements}
%If you'd like to thank anyone, place your comments here
%and remove the percent signs.
%\end{acknowledgements}

% BibTeX users please use one of
%\bibliographystyle{spbasic}      % basic style, author-year citations
%\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
%\bibliography{}   % name your BibTeX data base

% Non-BibTeX users please use
%\begin{thebibliography}{}
%
% and use \bibitem to create references. Consult the Instructions
% for authors for reference list style.
%
%\bibitem{RefJ}
% Format for Journal Reference
%Author, Article title, Journal, Volume, page numbers (year)
% Format for books
%\bibitem{RefB}
%Author, Book title, page numbers. Publisher, place (year)
% etc
%\end{thebibliography}

\end{document}
% end of file template.tex